---
title: "Gates Foundation Grants"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

# Purpose

This page documents Rick Gilmore's explorations of the Gates Foundation grants database.

# Set-up

```{r}
library(tidyverse)
```

# Download data

The grants data may be found here:

<https://www.gatesfoundation.org/about/committed-grants?topic=MNCH%20Discovery%20and%20Tools&yearAwardedEnd=2021&yearAwardedStart=2011>

```{r}
download.file(url = "https://www.gatesfoundation.org/-/media/files/bmgf-grants.csv",
              destfile = "csv/bmgr-grants.csv")
```

# Import & clean

```{r}
gates_txt <- readLines("csv/bmgr-grants.csv")

# Drop first line with update date
gates_csv <- gates_txt[2:length(gates_txt)]
writeLines(gates_csv, "csv/bmgr-grants-trim.csv")
```

Reimport.

```{r}
gates_df <- readr::read_csv("csv/bmgr-grants-trim.csv")
```

# Explore

Let's see what divisions are represented.

```{r}
unique(gates_df$DIVISION)
```
I'm specifically looking for "MNCH Discovery and Tools". This is a `TOPIC`.

```{r}
sum(stringr::str_detect(gates_df$TOPIC, "MNCH"))
```

## Filter on `MNCH`

```{r}
mnch_dt <- gates_df %>%
  dplyr::filter(., stringr::str_detect(TOPIC, "MNCH"))

str(mnch_dt)
```

```{r}
mnch_dt %>%
  ggplot(.) +
  aes(`DURATION (MONTHS)`) +
  geom_histogram()
```

```{r}
mnch_dt %>%
  ggplot(.) +
  aes(`AMOUNT COMMITTED`) +
  geom_histogram(bins = 20)
```

```{r}
mnch_dt <- mnch_dt %>%
  dplyr::select(
    .,
    GRANTEE,
    DIVISION,
    `GRANTEE COUNTRY`,
    `GRANTEE`,
    `DATE COMMITTED`,
    `DURATION (MONTHS)`,
    `AMOUNT COMMITTED`
  ) %>%
  dplyr::mutate(
    .,
    dollars_per_yr = `AMOUNT COMMITTED` / `DURATION (MONTHS)` * 12,
    year_committed = stringr::str_sub(`DATE COMMITTED`, start = 1, end = 4)
  ) %>%
  dplyr::arrange(.,
                 desc(year_committed),
                 `GRANTEE COUNTRY`,
                 `GRANTEE`,
                 desc(`AMOUNT COMMITTED`))


knitr::kable(mnch_dt, 'html')
```

Histogram of dollars-committed/project month.

```{r}
mnch_dt %>%
  ggplot(.) +
  aes(dollars_per_yr) +
  geom_histogram()
```

## Commitments per year

```{r}
mnch_dt_py <- mnch_dt %>%
  dplyr::group_by(., year_committed) %>%
  dplyr::summarize(., dollars_committed = sum(`AMOUNT COMMITTED`),
                   n_awards = n(),
                   avg_award = dollars_committed/n_awards)

mnch_dt_py %>%
  knitr::kable(., 'html')
```

Time series plots.

```{r}
mnch_dt_py %>%
  ggplot(.) +
  aes(year_committed, dollars_committed) +
  geom_point()
```

```{r}
mnch_dt_py %>%
  ggplot(.) +
  aes(year_committed, n_awards) +
  geom_point()
```

```{r}
mnch_dt_py %>%
  ggplot(.) +
  aes(year_committed, avg_award) +
  geom_point()
```

## Total commitments since 2011

```{r}
format(sum(mnch_dt_py$dollars_committed), digits = 12, nsmall = 0, big.mark = ",")
```

